---
title: "EDA wordcloud"
author: "UBC-MDS Team"
date: "20/05/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load libraries
library(tm)
library(wordcloud)
library(dplyr)
library(tidytext)
library(igraph)
library(ggraph)
library(reticulate)
```

These two wordcloud plots avoid `stopwords`. Additionally, in by-word analysis it avoids the word `staff` and for bigrams it avoids `would like`.

Warning comment: this file take a while to `knit` becasue it uses the preprocess.py **python** script and after transfroms the results to **R**, to follow with the process.

## By Word

### Wordcloud - question 1
```{r loading, include=FALSE}
# read data
wordcloud_q1 <- readxl::read_excel("../../data/interim/question1_models/ministries_Q1.xlsx")
wordcloud_q2 <- readxl::read_excel("../../data/interim/question2_models/ministries_Q2.xlsx")
```


```{r wordcloud, echo=FALSE, fig.width=5, fig.height=5, warning=FALSE}
# wordcloud code
only_comments <- wordcloud_q1 %>% select(Comment)

single_tokens <-
          only_comments %>% unnest_tokens(word, Comment)

words_df_q1 <- single_tokens %>%
          filter(word != "staff") %>% 
          count(word) %>%
          anti_join(stop_words, by = "word")

set.seed(42) # for reproducibility
wordcloud(words = words_df_q1$word, freq = words_df_q1$n, min.freq = 1,
          max.words=100, random.order=FALSE, rot.per=0.35,
          colors=brewer.pal(8, "Dark2"))
```


### Wordcloud - question 2 (Unsupervised dataset: 2015, 2018, 2020)
```{r wordcloud question 2, fig.width=5, fig.height=5,echo=FALSE, warning=FALSE}
# wordcloud code
only_comments <- wordcloud_q2 %>% select(Comment)# %>% anti_join(stop_words)

single_tokens <-
          only_comments %>% unnest_tokens(word, Comment)

words_df_q2 <- single_tokens %>% 
          filter(word != "staff") %>% 
          count(word) %>% 
          anti_join(stop_words, by = "word") 

set.seed(42) # for reproducibility
wordcloud(words = words_df_q2$word, freq = words_df_q2$n, min.freq = 1,
          max.words=100, random.order=FALSE, rot.per=0.35,
          colors=brewer.pal(8, "Dark2"))
```


## n-grams^[Note: For the bigram's wordcloud we use the preprocessed comments, but we didn't remove any additional words.]

```{python preprocess, echo=FALSE}
import sys
sys.path.append('../../src/data/')
from preprocess import Preprocessing

comment_q1_py = r["wordcloud_q1['Comment']"]
comment_q2_py = r["wordcloud_q2['Comment']"]

pp_q1_py = Preprocessing().general(comment_q1_py['Comment'])
pp_q2_py = Preprocessing().general(comment_q2_py['Comment'])
```

```{python, include=FALSE}
# # To run in python we could use:
# pp_q1_py
```

```{r, include=FALSE}
# # To run in R we can use:
# py$pp_q1_py

# pass the preprocessed data from python to R, for both questions
pp_q1_r <- data.frame(Comment = py$pp_q1_py)
df_q1 <- tibble::rowid_to_column(pp_q1_r, "ID")

pp_q2_r <- data.frame(Comment = py$pp_q2_py)
df_q2 <- tibble::rowid_to_column(pp_q2_r, "ID")

```


### Wordcloud of biagrams - question 1


```{r bigrams question 1, fig.width=5, fig.height=5, echo=FALSE, warning=FALSE}
# making the bigrams
bigrams_q1 <- df_q1 %>% 
  unnest_tokens(bigram, Comment, token ="ngrams", n=2) %>% 
  filter(bigram != "would like") %>%
  count(bigram, sort = TRUE)

# wordcloud code
pal=brewer.pal(8,"Blues")
pal=pal[-(1:3)]
wordcloud(bigrams_q1$bigram, bigrams_q1$n, max.words=75,random.order = F, colors=brewer.pal(8, "Dark2"))
```

### Wordcloud of biagrams - question 2


```{r bigrams question 2, fig.width=5, fig.height=5, echo=FALSE, warning=FALSE}
# making the bigrams
bigrams_q2 <- df_q2 %>%
  unnest_tokens(bigram, Comment, token ="ngrams", n=2) %>% 
  count(bigram, sort = TRUE)

# wordcloud code
pal=brewer.pal(8,"Blues")
pal=pal[-(1:3)]
wordcloud(bigrams_q2$bigram, bigrams_q2$n, max.words=75,random.order = F, colors=brewer.pal(8, "Dark2"))
```
